### 2019-01-27 1/exams/f.tex
final.

\begin{center}
  {\Large {\bf CMSC 131A, Final Exam}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2017 \ \\
  }
\end{center}

@@

\begin{verbatim}
(define S 1)
(define (q z)
  (cond [(< 4 z) (+ z S)]
        [else 9]))
(q 8)
\end{verbatim}

@@

(q 8)                                --[plug]-->
^^^^^
(cond [(< 4 8) (+ 8 S)] [else 9])    --[arith]-->
       ^^^^^^^
(cond [#true (+ 8 S)] [else 9])      --[cond]-->
^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
(+ 8 S)                              --[const]-->
     ^
(+ 8 1)                              --[arith]-->
^^^^^^^
9
\end{SaveVerbatim}

@@

\begin{verbatim}
(define-struct qt (nw ne se sw))
\end{verbatim}

@@

- make-qt  : Constructor
- qt-nw    : Accessor
- qt-ne    : Accessor
- qt-se    : Accessor
- qt-sw    : Accessor
- qt?      : Predicate
\end{SaveVerbatim}

@@

\noindent
{\bf Problem 3 (15 points).}
%
Consider the following data definition (and examples) for representing sentences:
\begin{verbatim}
;; A Sentence is a [NEListof String]

(define sent1 (list "a" "dog" "barked"))
(define sent2 (list "a" "cat" "meowed"))
\end{verbatim}
Recall the following data definition of {\tt [NEListof X]}:
\begin{verbatim}
;; A [NEListof X] is one of:
;; - (cons X '())
;; - (cons X [NEListof X])
\end{verbatim}
Given a {\tt Sentence}, a \emph{bigram} is a list of two consecutive
elements of the sentence. For example, {\tt sent1} has exactly two
bigrams:
\begin{verbatim}
(list "a" "dog")
(list "dog" "barked")
\end{verbatim}
Using the following data definition,
\begin{verbatim}
;; a Bigram is a (list String String)
\end{verbatim}
design a function {\tt bigrams} which takes a {\tt Sentence} and
returns a list of {\tt Bigrams}.

@@

;; bigrams : Sentence -> [Listof Bigram]
;; List of all adjacent words in a sentence
(check-expect (bigrams (list "one")) '())
(check-expect (bigrams sent1)
              (list (list "a" "dog") (list "dog" "barked")))
(define (bigrams sent)
  (cond [(empty? (rest sent)) '()]
        [else
         (cons (list (first sent) (second sent))
               (bigrams (rest sent)))]))
\end{SaveVerbatim}

@@

\noindent
{\bf Problem 4 (15 points).}
%
Here is a program that, given a sentence (see problem 3), computes a
two element list consisting of the minimal and maximal words according
to alphabetic ordering ({\tt string<?}).

\begin{verbatim}
;; min-and-max : Sentence -> (list String String)
;; Produces the min and max words in the sentence.
(check-expect (min-and-max (list "the" "zoo" "has" "aardvarks"))
              (list "aardvarks" "zoo"))
(define (min-and-max s)
  (cond [(empty? (rest s)) (list (first s) (first s))]
        [else
         (local [(define w (first s))
                 (define m+m (min-and-max (rest s)))]
            (cond [(string<? w (first m+m))
                   (list w (second m+m))]
                  [(string<? (second m+m) w)
                   (list (first m+m) w)]
                  [else m+m]))]))
\end{verbatim}
Redesign this program to use an accumulator-based helper function with
two accumulators: the mininmal and maximal elements of the sentence
seen so far.

@@

(define (min-and-max s)
  (min-and-max/a (rest s) (first s) (first s)))

;; min-and-max/a : [Listof String] String String -> (list String String)
;; Accumulator: min and max element seen so far
(define (min-and-max/a s min max)
  (cond [(empty? s) (list min max)]
        [(cons? s)
         (local [(define w (first s))]
           (cond [(string<? w min) 
                  (min-and-max/a (rest s) w max)]
                 [(string<? max w)
                  (min-and-max/a (rest s) min w)]
                 [else
                  (min-and-max/a (rest s) min max)]))]))
\end{SaveVerbatim}

@@

[Space for problem 4.]

\newpage 
\noindent
{\bf Problem 5 (15 points).}
%
Design the following program for determining if two sentences (see problem 3) are equal,
meaning they are lists of the same length with the same elements (in the same order):
\begin{verbatim}
;; sent=? : Sentence Sentence -> Boolean
;; Are the two sentences the same?
\end{verbatim}
(You may not use {\tt equal?} in the design of this program.)

@@

;; sent=? : Sentence Sentence -> Boolean
;; Are the two sentences equal?
(check-expect (sent=? (list "one") (list "one")) #true)
(check-expect (sent=? (list "one") (list "two")) #false)
(check-expect (sent=? (list "one") (list "one" "two")) #false)
(check-expect (sent=? (list "one" "two") (list "one" "two")) #true)
(define (sent=? s1 s2)
  (cond [(empty? (rest s1))
         (and (empty? (rest s2))
              (string=? (first s1) (first s2)))]
        [else
         (and (not (empty? (rest s2)))
              (string=? (first s1) (first s2))
              (sent=? (rest s1) (rest s2)))]))
\end{SaveVerbatim}

\newpage
\noindent
{\bf Problem 6 (15 points).}  
%
During your internship at the University of Maryland's Center for
Bioinformatics and Computational Biology (CBCB), you're asked to
design a program for computing the \emph{reverse complement} of DNA strands.
%
A DNA strand is an arbitrarily long sequence of bases, which are
either: C, G, A, or T.  A and T are complements of each other, as are
C and G.  The reverse complement of a DNA strand is formed by
reversing the strand and taking the complement of each symbol.  For
example, the reverse complement of AAAACCCGGT is ACCGGGTTTT.
%
Design a program that, given a DNA strand, computes its reverse complement.

;; A DNA is a [Listof Base]
;; A Base is one of: "A", "C", "G", "T"
;; Interp: DNA is a sequence of nucleotide bases

;; comp : Base -> Base
;; Compute the complement of the given base
(check-expect (comp "A") "T")
(check-expect (comp "T") "A")
(check-expect (comp "C") "G")
(check-expect (comp "G") "C")
(define (comp b)
  (cond [(string=? b "A") "T"]
        [(string=? b "T") "A"]
        [(string=? b "C") "G"]
        [(string=? b "G") "C"]))
\end{SaveVerbatim}

\noindent
{\bf Problem 7 (15 points).} Design a program that takes two
strings and counts how many times the first string occurs within the
second.
For example,
\begin{itemize}
\item {\tt "ho"} occurs three times in {\tt "ho ho ho"}, while
\item {\tt "ho ho"} occurs once in {\tt "ho ho ho"} (notice that occurences
  cannot overlap).
\end{itemize}
You may assume the first string is non-empty.
%
It may be helpful to use the {\tt substring} function.  Recall: {\tt
  (substring "hello world" 1 5)} produces {\tt "ello"} and {\tt
  (substring "hello world" 4)} produces {\tt "o world"}.

;; occurs : String String -> Natural
;; Count (non-overlapping) occurences of first string in second
;; Assume: first string is non-empty
;; Termination: because s1 is non-empty, each recursive call is on a strictly
;; smaller s2, therefore eventually reaching first cond clause (trivial case).
(check-expect (occurs "ho" "ho ho ho") 3)
(check-expect (occurs "ho ho" "ho ho ho") 1)
(define (occurs s1 s2)
  (local [(define s1-len (string-length s1))]
    (cond [(< (string-length s2) s1-len) 0]
          [(string=? s1 (substring s2 0 s1-len))
           (add1 (occurs s1 (substring s2 s1-len)))]
          [else
           (occurs s1 (substring s2 1))])))
\end{SaveVerbatim}

>>>>

\begin{center}
  {\Large {\bf CMSC 131A, Final Exam}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2018 \ \\
  }
\end{center}

@@

\begin{verbatim}
(define S 1)
(define (w p) (< S p))

(cond [(w 9) (* 8 2)]
      [else 9])
\end{verbatim}

@@

(cond [(w 9) (* 8 2)] [else 9])      --[plug]-->
       ^^^^^
(cond [(< S 9) (* 8 2)] [else 9])    --[const]-->
          ^
(cond [(< 1 9) (* 8 2)] [else 9])    --[arith]-->
       ^^^^^^^
(cond [#true (* 8 2)] [else 9])      --[cond]-->
^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
(* 8 2)                              --[arith]-->
^^^^^^^
16
\end{SaveVerbatim}

@@

\begin{verbatim}
(define-struct roger (waters rabbit))
\end{verbatim}

@@

- make-roger   : Constructor
- roger-waters : Accessor
- roger-rabbit : Accessor
- roger?       : Predicate
\end{SaveVerbatim}

@@

\noindent
{\bf Problem 3 (15 points).}
%
Imagine designing the following two functions.  Now design an
abstraction of both and define each original function in terms of that
abstraction.  (It's not necessary to give the ``unabstracted''
definitions unless it's helpful to you.)
\begin{verbatim}
;; find-even : [Listof Number] -> [Maybe Number]
;; Find first even number, if there is one (or #false)
(check-expect (find-even '(1 2 3 4)) 2)
(check-expect (find-even '(1 3 5)) #false)

;; find-long : [Listof String] -> [Maybe String]
;; Find first string longer than 5 letters, if there is one (or #false)
(check-expect (find-long '("a" "really" "long")) "really")
(check-expect (find-long '("a" "b" "c")) #false)
\end{verbatim}
Recall that {\tt [Maybe X]} is defined as:
\begin{verbatim}
;; A [Maybe X] is one of:
;; - X
;; - #false
\end{verbatim}

@@

;; find : [Listof X] [X -> Boolean] -> [Maybe X]
;; Find first element satisfying the predicate, if there is one (or #false)
(define (find lon pred)
  (cond [(empty? lon) #false]
        [(cons? lon)
         (cond [(pred (first lon)) (first lon)]
               [else (find-even (rest lon))])]))

(define (find-even lon)
  (find lon even?))

(define (find-long los)
  (find los (lambda (s) (> (string-length s) 5))))
\end{SaveVerbatim}

@@

[Space for problem 3.]

\noindent
{\bf Problem 4 (15 points).}
%
Here is a data definition for binary trees:
\begin{verbatim}
;; A [BT X] is one of:
;; - (make-leaf)
;; - (make-node X [BT X] [BT X])
(define-struct node (val left right))
(define-struct leaf ())
\end{verbatim}
Design a function that consumes a binary tree of strings and
constructs a single string that results from concating the strings
in left to val to right order.  For example:
\begin{verbatim}
(define l (make-leaf))
(define t (make-node "b" (make-node "a" l l) (make-node "c" l l)))
\end{verbatim}
The function should produce {\tt "abc"} on {\tt t}.

@@

\end{SaveVerbatim}

%% Using an accumulator-based design, design the program {\tt to10} that
%% converts a list of digits into its correspdonding number.  The
%% elements are listed in order of most to least significance,
%% so for example, {\tt (to10 '(9 1 2 3))} should produce 9123.
%% \begin{verbatim}
%% ;; A Digit is a Natural in [0,9]

%% ;; to10 : [Listof Digit] -> Number
%% ;; Convert digits (in order of most to least significance) to a number
%% (check-expect (to10 '(1 2 3)) 123)
%% (check-expect (to10 '(7 2 8)) 728)
%% \end{verbatim}
%% Here is a hint, 728 can be computed as {\tt (+ 8 (* 10 (+ 2 (* 10 (+ 7 0)))))}.
%% A design that does not use an accumulator will receive only partial credit.

%% \begin{SaveVerbatim}{VerbEnv}

%% SOLUTION:

%% (define (to10 lod)
%%   (local [(define (to10/a lod a)
%%             (cond [(empty? lod) a]
%%                   [(cons? lod)
%%                    (to10/a (rest lod)
%%                            (+ (first lod)
%%                               (* 10 a)))]))]                              
%%     (to10/a lod 0)))
%% \end{SaveVerbatim}

%% \censor{%
%% \noindent
%% \BUseVerbatim{VerbEnv}}

\newpage 
\noindent
{\bf Problem 5 (15 points).}
%
Design the following variant of {\tt map} that takes a binary function
and \emph{two} lists that are assumed to be of equal length and
produces list of results formed by applying the function to the
elements of each list.  Only tests are given, you have to come up with
the rest:
\begin{verbatim}
(check-expect (2map < '(1 4 3) '(3 1 2)) '(#true #false #false))
(check-expect (2map string-ith '("fred" "wilma") '(1 0)) '("r" "w"))
\end{verbatim}

;; 2map : [X Y -> Z] [Listof X] [Listof Y] -> [Listof Z]
;; Binary variant of map
;; ASSUME: lists are equal length
(define (2map f xs ys)
  (cond [(empty? xs) '()]
        [(cons? xs)
         (cons (f (first xs) (first ys))
               (2map f (rest xs) (rest ys)))]))
\end{SaveVerbatim}

\newpage
\noindent
{\bf Problem 6 (15 points).}  
%
During your internship at the University of Maryland's Center for
Bioinformatics and Computational Biology (CBCB), you're asked to
improve code written by last year's intern.  That intern designed a
program for computing the \emph{reverse complement} of DNA strands.
%
A DNA strand is an arbitrarily long sequence of bases, which are
either: C, G, A, or T.  A and T are complements of each other, as are
C and G.  The reverse complement of a DNA strand is formed by
reversing the strand and taking the complement of each symbol.  For
example, the reverse complement of AAAACCCGGT is ACCGGGTTTT.
%
Last year's intern came up with the following:
\begin{verbatim}
;; A DNA is a [Listof Base]
;; A Base is one of: "A", "C", "G", "T"
;; Interp: DNA is a sequence of nucleotide bases

;; comp : Base -> Base
;; Compute the complement of the given base
(check-expect (comp "A") "T")
(check-expect (comp "T") "A")
(check-expect (comp "C") "G")
(check-expect (comp "G") "C")
(define (comp b)
  (cond [(string=? b "A") "T"]
        [(string=? b "T") "A"]
        [(string=? b "C") "G"]
        [(string=? b "G") "C"]))
\end{verbatim}
You notice that {\tt rev-comp} traverses the DNA sequence twice, once
with {\tt map} and once with {\tt reverse}.  Improve this design by
revising {\tt rev-comp} to use an accumulator to traverse the DNA
sequence just once.

\noindent
[Space for problem 6.]

(define (rev-comp dna)
  (local [;; DNA DNA -> DNA
          ;; ACCUM: the reverse complement of the sequence so far
          (define (rev-comp/a dna rcomp)
            (cond [(empty? dna) rcomp]
                  [(cons? dna)
                   (rev-comp/a (rest dna)
                               (cons (comp (first dna)) rcomp))]))]
     (rev-comp/a dna '())))                      
\end{SaveVerbatim}

@@

\noindent
{\bf Problem 7 (15 points).} Design a program that takes two
strings and counts how many times the first string occurs within the
second, {\bf including possibly overlapping occurences}.
For example,
\begin{itemize}
\item {\tt "ho"} occurs three times in {\tt "ho ho ho"}, while
\item {\tt "ho ho"} occurs {\bf twice} in {\tt "ho ho ho"} (notice that occurences
  overlap on the middle {\tt "ho"}).
\end{itemize}
You may assume the first string is non-empty.
%
It may be helpful to use the {\tt substring} function.  Recall: {\tt
  (substring "hello world" 1 5)} produces {\tt "ello"} and {\tt
  (substring "hello world" 4)} produces {\tt "o world"}.

@@

;; occurs : String String -> Natural
;; Count (possibly overlapping) occurences of first string in second
;; Assume: first string is non-empty
;; Termination: each recursive call is on a strictly
;; smaller s2, therefore eventually reaching first cond clause (trivial case).
(check-expect (occurs "ho" "ho ho ho") 3)
(check-expect (occurs "ho ho" "ho ho ho") 2)
(define (occurs s1 s2)
  (local [(define s1-len (string-length s1))]
    (cond [(< (string-length s2) s1-len) 0]
          [(string=? s1 (substring s2 0 s1-len))
           (add1 (occurs s1 (substring s2 1)))]
          [else
           (occurs s1 (substring s2 1))])))
\end{SaveVerbatim}
